Dependent Object Types

### 2012-08-30 doc/candidacy/report.tex
just copied the context-free syntax comment from nuObj. Requires editing

\input{macros}

@@

>>>>

\input{macros}
\input{nuobj}

@@

%% TODO beg edit
A {\em term} denotes an object or a class. It can be of the following
five forms.
\begin{itemize}
\item A {\em simple name} $x$, which denotes an object.
\item A {\em selection} $t.l$, which can denote either an object or a class.
\item An {\em object creation} \mbox{$\newin{x \larrowc t}{u}$},
which defines a fresh instance $x$ of class $t$.
The scope of this object is the term $u$.
\item 
A {\em class template} $\cls{x: S \sep \tup d}$ where $\tup d$ is a
sequence of {\em definitions} which associate term labels with values
and type labels with types. This acts as a template to construct
objects with the members defined by the definitions.  The name $x$ of
type $S$ stands for ``self'', i.e.\ the object being constructed from
the template.  Its scope is the definition sequence $\tup d$.  A term
or type can refer via $x.l$ \comment{or \mbox{$x.\type \Tsel L$}} to some other
member of that object. No textual sequence constraint applies to such
references; in particular it is possible that a binding refers to
itself or to bindings defined later in the same record.  This
distinguishes our type system from earlier type systems for records
\cite{cardelli-mitchell:records} or modules \cite{harper94type}.
\item 
A {\em mixin composition} $t \withsub{S} u$, which forms a combined
class from the two classes to which $t$ and $u$ evaluate. Here, $S$
is the type of ``self'' in the combined class. 
\end{itemize}
A {\em value} is a simple name or a class template.
A {\em path} $p$ is a name $x$ followed
by a possibly empty sequence of selections, e.g.\ $x . l_1 .; \ldots;
. l_n$.

The syntax of {\em types} in our system closely follows the syntax of
 terms.  A type can be of the following five forms.
\begin{itemize}
\item A {\em singleton type} $p.\type$. This type represents the set of
values which has as only element the object referenced by the path
$p$. Singleton types are the only way a type can depend on a term in
$\NTO$.
\item 
A {\em type selection} $T\Tsel L$, which represents the type component labelled $L$ of type $T$.
\item
A {\em record type} $\obj{x \sep \tup D}$ where $\tup D$ is a sequence
of {\em declarations} which can be value bindings or type bindings.  A
{\em value binding} $l: T$ associates a term label $l$ with its type
$T$.  {\em Type bindings} come in three different forms:

First, the binding $L = T$ defines $L$ to be an {\em alias} for $T$. 
Second, the binding $L \expand T$ defines $L$ to be a {\em new type} which
{\em expands} to type $T$. That is, $L$ is a subtype of $T$ which has
exactly the members defined by $T$; furthermore, one can create
objects of type $L$ from a class which defines all members of $T$.
Third, the binding $L <: T$ defines $L$ to be an {\em abstract type}
which is known to be a subtype of its bound $T$.

We let the meta-variable $\preceq$ range over $=$ and $\expand$, and
let $\tbind$ range over $=$, $\expand$, and $<:$.  The name $x$ stands
for ``self''; its type is assumed to be the record type itself. We let
the letter $R$ range over record types.
\item
A {\em compound type} $T \with U$. This type contains all members of
types $T$ and $U$. The subtyping relation for compound types is the
same as the one for intersection types \cite{barendregt-coppo:filter},
but the formation rules are more restrictive.  Where $T$ and $U$ have
a member with the same label, the compound type contains the member
defined in $U$. That member definition must be more specific
(see Section~\ref{sec:type-system}) than the corresponding member
definition in $T$.
\item
A {\em class type} $\cls{x: S \sep \tup D}$, which contains as values
classes that instantiate to objects of type $\obj{x \sep \tup D}$, or
some subtype of it.  $x$ is again the name for ``self''.  It now comes
with an explicit type $S$ which may be different from $\obj{x \sep
\tup D}$. Definitions in $S$ which are missing from $\tup D$ play the
role of abstract members. Such members can be referred to from other
definitions in the class, but they are not defined in the class
itself. Instead, these members must be defined in other classes which
are composed with the class itself in a mixin composition.
Definitions which are present in $\tup D$ but missing in $S$ play in
some sense the role of non-virtual members -- they are not referred to
via ``self'' from inside the class, so overriding them does not change
existing behavior. Definitions present in both $S$ and $\tup D$ play
the role of virtual members.
\end{itemize}
%% TODO end edit
